Verification of Stencil Computations with Spatial Specifications

### 2016-07-14 paper.tex
submitted version of the paper
Modify submitted version to add DO NOT DISTRIBUTE in title banner and footer
added missing last line to navier example
update region declaration syntax in the paper

\title{Verification of Stencil Computations with Spatial
  Specifications}
\authorinfo{}{}{}
%\authorinfo{Dominic Orchard \and Mistral Contrastin
%\and Matthew Danish \and Andrew Rice}{University of Cambridge}{}

@@

laplu = (u(i+1,j)-2.0*u(i,j)+u(i-1,j))/delx/delx+ &
  (u(i,j+1)-2.0*u(i,j)+u(i,j-1))/dely/dely
\end{minted}
(a) Excerpt of a Navier-Stokes fluid simulator
(based on~\citet{griebel1997numerical}) showing highly-detailed stencil computation.
\vspace{0.5em}
\hrule
%
\begin{minted}[linenos=false,xleftmargin=1em]{fortran}
!= stencil centered(depth=1,dim=1)*reflexive(dim=2) + centered(depth=1,dim=2)*reflexive(dim=1) :: u

@@

\emph{Regions} are the central building blocks of spatial
specifications. Regions can either be declared alone with a
\nonterm{regionDec}, assigning a region specification \nonterm{region} to
a region variable \nonterm{rvar} or given directly within a
\nonterm{spatial} specification.

@@

\begin{figure}[t]
\vspace{-0.5em}
\begin{align*}
\def\arraystretch{1.2}
\setlength{\arraycolsep}{0.2em}
\newcommand{\dimTy}{\mathbb{N}_{>0}}
\begin{array}{rl}
\nonterm{specification} ::= & \nonterm{regionDec} \mid \nonterm{specDec} \\
\nonterm{specDec} ::= & \term{stencil} \; \nonterm{spec} \;
                        \texttt{::} \; v \\
\nonterm{regionDec} ::= &  \texttt{region} \; \nonterm{rvar} \; \texttt{=} \;
                         \nonterm{region}\\[0.4em]
%\nonterm{spec} ::= & \nonterm{spatial} \mid \nonterm{temporal}
%\\[1em]
\nonterm{spec} ::= & [\nonterm{approx},] \; [\nonterm{mult},] \; \nonterm{region} \\
\nonterm{mult} ::= & \term{readOnce} \\
\nonterm{approx} ::= & \term{atMost} \; \mid \; \term{atLeast} \\[0.1em]
\nonterm{region} ::= & \nonterm{rvar} \; \\
\multicolumn{2}{l}{\qquad\qquad \mid \; \stenRefl{\dimTy}} \\
\multicolumn{2}{l}{\qquad\qquad \mid \;\stenFwd{\mathbb{N}_{>0}}{\dimTy}{\;[, \texttt{irreflexive}]}} \\
\multicolumn{2}{l}{\qquad\qquad \mid \; \stenBwd{\mathbb{N}_{>0}}{\dimTy}{\;[, \texttt{irreflexive}]}} \\
\multicolumn{2}{l}{\qquad\qquad \mid \; \stenCen{\mathbb{N}_{>0}}{\dimTy}{\;[, \texttt{irreflexive}]}} \\
\multicolumn{2}{l}{\qquad\qquad \mid \; \nonterm{region} \; \term{+}
  \; \nonterm{region} \; \mid \; \nonterm{region} \; \term{*} \; \nonterm{region}} \\[0.5em]
%\multicolumn{2}{l}{\qquad\qquad \mid \; \nonterm{rvar}}  
%\\[0.5em]
%\nonterm{temporal} ::= \; & \term{dependency} \; (v \; \{ , v \}) [, \texttt{mutual}]
%  \\[0.5em]
\nonterm{rvar} ::= \; & [\text{\term{a}-\term{z}$\,$\term{A}-\term{Z}$\,$\term{0}-\term{9}}]+\\[-0.5em]
\end{array}
\end{align*}
\caption{Specification syntax (EBNF grammar)}
\label{fig:syntax}
\vspace{-0.8em}
\end{figure}

@@

Region specifications can be assigned to region variables via
a region declaration, which can be used later to form a spatial
specification. For example, the specification from the previous
 example can be restated as:
%%
\begin{minted}{fortran}
!= region r1 = forward(depth=1, dim=1)
!= region r2 = forward(depth=1, dim=2)
!= region robertsCross = r1*r2
!= stencil robertsCross :: a
\end{minted}
This is especially useful for common stencil patterns, such as Roberts cross,
as the region can be defined once and reused.
%%
\paragraph{Modifiers}
%%
Region specifications can be modified 
by \emph{approximation} and \emph{multiplicity} information 
(in the \textit{spec} rule of Fig.~\ref{fig:syntax}).

@@

\begin{example}
We give an extended example of verifying a stencil computation against
its specification. We consider a more general, but still consistent, version of the \textit{five-point}
stencil specification inferred previously in
Section~\ref{subsec:inference}:
%
\begin{minted}[breakindent=8.1em]{fortran}
!= region fivepoint = centered(depth=1,dim=1) + centered(depth=1,dim=2)
!= stencil readOnce, fivepoint :: b
a(i,j) = b(i,j-1)+b(i,j)+b(i,j+1)+b(i-1,j)+b(i+1,j)
\end{minted}
%
We refer to the assignment on line three as $l3$ for brevity.

>>>>

\title{Verification of Stencil Computations with Spatial
  Specifications}
\titlebanner{DRAFT -- Do not distribute}
%\authorinfo{}{}{}
\authorinfo{Dominic Orchard \and Mistral Contrastin
\and Matthew Danish \and Andrew Rice}{University of Cambridge}{}

\preprintfooter{DO NOT DISTRIBUTE}

@@

laplu = (u(i+1,j)-2.0*u(i,j)+u(i-1,j))/delx/delx+ &
  (u(i,j+1)-2.0*u(i,j)+u(i,j-1))/dely/dely

f(i,j) = u(i,j)+del_t*(laplu/Re-du2dx-duvdy)
\end{minted}
(a) Excerpt of a Navier-Stokes fluid simulator
(based on~\citet{griebel1997numerical}) showing highly-detailed stencil computation.
\vspace{0.5em}
\hrule
%
\begin{minted}[linenos=false,xleftmargin=1em]{fortran}
!= stencil centered(depth=1,dim=1)*reflexive(dim=2) + centered(depth=1,dim=2)*reflexive(dim=1) :: u

@@

\emph{Regions} are the central building blocks of spatial
specifications. Regions can either be declared alone with a
\nonterm{regionDec}, assigning a region specification \nonterm{region} to
a region variable \nonterm{rvar} (via $\texttt{::}$) or given directly within a
\nonterm{spatial} specification.

@@

\begin{figure}[t]
\vspace{-0.5em}
\begin{align*}
\def\arraystretch{1.2}
\setlength{\arraycolsep}{0.2em}
\newcommand{\dimTy}{\mathbb{N}_{>0}}
\begin{array}{rl}
\nonterm{specification} ::= & \nonterm{regionDec} \mid \nonterm{specDec} \\
\nonterm{specDec} ::= & \term{stencil} \; \nonterm{spec} \;
                        \texttt{::} \; v \\
\nonterm{regionDec} ::= &  \texttt{region} \; \texttt{::} \; \nonterm{rvar} \; \texttt{=} \;
                         \nonterm{region}\\[0.4em]
%\nonterm{spec} ::= & \nonterm{spatial} \mid \nonterm{temporal}
%\\[1em]
\nonterm{spec} ::= & [\nonterm{approx},] \; [\nonterm{mult},] \; \nonterm{region} \\
\nonterm{mult} ::= & \term{readOnce} \\
\nonterm{approx} ::= & \term{atMost} \; \mid \; \term{atLeast} \\[0.1em]
\nonterm{region} ::= & \nonterm{rvar} \; \\
\multicolumn{2}{l}{\qquad\qquad \mid \; \stenRefl{\dimTy}} \\
\multicolumn{2}{l}{\qquad\qquad \mid \;\stenFwd{\mathbb{N}_{>0}}{\dimTy}{\;[, \texttt{irreflexive}]}} \\
\multicolumn{2}{l}{\qquad\qquad \mid \; \stenBwd{\mathbb{N}_{>0}}{\dimTy}{\;[, \texttt{irreflexive}]}} \\
\multicolumn{2}{l}{\qquad\qquad \mid \; \stenCen{\mathbb{N}_{>0}}{\dimTy}{\;[, \texttt{irreflexive}]}} \\
\multicolumn{2}{l}{\qquad\qquad \mid \; \nonterm{region} \; \term{+}
  \; \nonterm{region} \; \mid \; \nonterm{region} \; \term{*} \; \nonterm{region}} \\[0.5em]
%\multicolumn{2}{l}{\qquad\qquad \mid \; \nonterm{rvar}}  
%\\[0.5em]
%\nonterm{temporal} ::= \; & \term{dependency} \; (v \; \{ , v \}) [, \texttt{mutual}]
%  \\[0.5em]
\nonterm{rvar} ::= \; & [\text{\term{a}-\term{z}$\,$\term{A}-\term{Z}$\,$\term{0}-\term{9}}]+\\[-0.5em]
\end{array}
\end{align*}
\caption{Specification syntax (EBNF grammar)}
\label{fig:syntax}
\vspace{-0.8em}
\end{figure}

@@

Region specifications can be assigned to region variables via
a region declaration, which can be used later to form a spatial
specification. For example, the specification from the previous
 example can be restated as:
%%
\begin{minted}{fortran}
!= region :: r1 = forward(depth=1, dim=1)
!= region :: r2 = forward(depth=1, dim=2)
!= region :: robertsCross = r1*r2
!= stencil robertsCross :: a
\end{minted}
This is especially useful for common stencil patterns, such as Roberts cross,
as the region can be defined once and reused.
%%
\paragraph{Modifiers}
%%
Region specifications can be modified 
by \emph{approximation} and \emph{multiplicity} information 
(in the \textit{spec} rule of Fig.~\ref{fig:syntax}).

@@

\begin{example}
We give an extended example of verifying a stencil computation against
its specification. We consider a more general, but still consistent, version of the \textit{five-point}
stencil specification inferred previously in
Section~\ref{subsec:inference}:
%
\begin{minted}[breakindent=8.1em]{fortran}
!= region :: fivepoint = centered(depth=1,dim=1) + centered(depth=1,dim=2)
!= stencil readOnce, fivepoint :: b
a(i,j) = b(i,j-1)+b(i,j)+b(i,j+1)+b(i-1,j)+b(i+1,j)
\end{minted}
%
We refer to the assignment on line three as $l3$ for brevity.
